Browsernative Microservices

### 2017-03-01 paper.tex
pdf/tex build

\subsection{2.1. Componentization via
Services}\label{componentization-via-services}

@@

\subsection{2.2. Organized around Business
Capabilities}\label{organized-around-business-capabilities}

@@

\subsection{2.4. Decentralized
Governance}\label{decentralized-governance}

@@

The spirit of freedom cannot be applied universally to
\emph{browsernative microservices} as the browser and its underlying DOM
will, to a certain degree, be the limiting factor. Talking about the
browser, a reader might be falsely tempted to narrow his perception
towards the obvious VIEW layer only. In recent years the major browser
engines have grown to fully-fledged app deployment platforms offering
connectors to build-in databases, multithreading support and
ever-growing JS build-ins like speech synthesis or push notifications.
So-called \textbf{Progressive Web Apps}\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}, a bunch of criteria for building good browser apps, can
achieve a similar look and feel like native apps. And last but not least
services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}
  \#\# 2.5. Decentralized Data Management} effectively compiling `the
web' to native machine code lower the boundary between native and
browser code even further.

@@

Firstly, Abramov defined a pattern around \textbf{presentational
components} only related with the concern about \emph{how things look}.
This component literally does not know anything about the service in
question which makes the component highly flexible and reusable. It is
controlled solely from the outside, receiving data and dispatching
unbiased events on user interaction.{[}20{]} Most probably every
presentational component embodies more HTML/CSS markup and less JS code.
It should encapsulate its styles from bleeding out and protect its
styles from being overwritten. Furthermore, it may contain several
templates to change its look on different demands.

Secondly, Abramov described components he refers to as
\textbf{containers}. A container component is concerned with \emph{how
things work}.{[}20{]} Containers act as invisible wrappers around
presentational components acting in the sense of UNIX filters. Their job
is to fetch data from child nodes, aggregating events, interacting with
the model and push state back to the presentational components.
Consequently they might contain more JS and less if any HTML markup. We
probably do not need to utilize shadow DOM as no styles are involved.

@@

The root container \texttt{\textless{}shop-checkout\textgreater{}} is
basically just an encapsulation layer in terms of service
functionalities. Encapsulation of CSS will not be an issue at this point
as no styling is involved. A simplified \emph{root container} for the
checkout might look like the following code snippet:

@@

The second layer of the checkout microservice like
\texttt{\textless{}sign-in\textgreater{}} or
\texttt{\textless{}shipping-details\textgreater{}} still contains mostly
logic and little or less styling. The job of a \emph{container
component} is to control its underlying presentational components and to
define a \textbf{set of ACTIONS towards the model}. Every
\emph{container} mounted directly under the service root may have a
dedicated area inside the worker thread reserved for its duties. For
example, a typical \texttt{\textless{}sign-in\textgreater{}} contains
merely two fields, username and password and a submit button. The
\emph{container component} waits for a submit action, aggregating the
credentials, and might add some semantics to them. Field values and the
action message will be dispatched towards the model for further
processing like initiating an authorization process.

@@

\section{5. Thinking further}\label{thinking-further}

In the search for an independent \textbf{browsernative} approach this
paper left out third-party libraries so far. Nevertheless, it should be
clear to this point that the example in the last section is far away
from being usable in production. At first, from a technical perspective
web components are relatively young and at the time of writing only
usable in \textbf{newer Chrome and Opera browsers}. Most probably it
will take a long time until native web components can be safely used
without additional legacy support. The troublesome \emph{HTML imports}
spec will certainly need even longer which inevitably opens up the
question if web components are useful from today onwards. And second, do
web components align with the notion of simplicity if too many downsides
and possible pitfalls require attention?!

Any profound discussion about web components would be incomplete without
touching component alternatives like \textbf{React}. In the last couple
years React introduced new concepts to web development like the idea of
state, passing data as properties and declarative event management just
to name a few. Those design decisions were made to simplify browser
development with full browser support right from the start. Overall,
React's perception about UI development created much of a hype around
the library and technologies like the virtual DOM while web components
never caught up in momentum at a comparable rate. According to Github,
Google's Polymer project is even six months older than React but has
less than one third of its stars. Other notable web component projects
like Bosonic\footnote{https://github.com/bosonic/bosonic/} or Mozilla's
X-tag\footnote{https://github.com/x-tag/core} are more dead than alive,
too.

It seems like the web developing community in a broader sense values
\textbf{standardized procedures, tooling and browser support} in
frameworks like React (and many others to be fair). Previously mentioned
web component libraries clearly missed a real selling point compared to
React or Angular. Providing guidelines, simpler syntax and features are
missed out or are too cumbersome to use at all. For the future it
remains unclear if web components eventually gain wider adoption within
the UI developing community or will be adopted as low-level technology
for framework development emphasized by Sebastian Markbage, one of the
React creators.{[}21{]}

Even though the drawbacks seems heavy weight, there are notable attempts
to bring web components into production. A very good example is the
``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
Following the functional rendering model from React, it combines native
methods with additional functionalities known from React like
declarative event management. Given the focus on native technologies it
only weights around 4kb (minified and gzipped) which is ten times less
than React. \textbf{Micro UI frameworks} like Skate could certainly be
the near future of web development, offering a real selling point.
Previously mentioned specs could lead to a modular, pluggable building
pipelines offering the same features and tooling as React already does.
After all, web components fit better with the microservice
\textbf{decentralization aspect} than `walled garden' frameworks.

>>>>

\subsection{2.1. Componentization via
services}\label{componentization-via-services}

@@

\subsection{2.2. Organized around business
capabilities}\label{organized-around-business-capabilities}

@@

\subsection{2.4. Decentralized
governance}\label{decentralized-governance}

@@

The spirit of freedom cannot be applied universally to
\emph{browsernative microservices} as the browser and its underlying DOM
will, to a certain degree, be the limiting factor. Talking about the
browser, a reader might be falsely tempted to narrow his perception
towards the obvious VIEW layer only. In recent years the major browser
engines have grown to fully-fledged app deployment platforms offering
connectors to build-in databases, multithreading support and
ever-growing JS build-ins like speech synthesis or push notifications.
So-called \textbf{Progressive Web Apps}\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}, a bunch of criteria for building good browser apps, can
achieve a similar look and feel like native apps. And last but not least
services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}
  \#\# 2.5. Decentralized data management} effectively compiling `the
web' to native machine code lower the boundary between native and
browser code even further.

@@

Firstly, Abramov defined a pattern around \textbf{presentational
components} only related with the concern about \textbf{how things
look}. This component literally does not know anything about the service
in question which makes the component highly flexible and reusable. It
is controlled solely from the outside, receiving data and dispatching
unbiased events on user interaction.{[}20{]} Most probably every
presentational component embodies more HTML/CSS markup and less JS code.
It should encapsulate its styles from bleeding out and protect its
styles from being overwritten. Furthermore, it may contain several
templates to change its look on different demands.

Secondly, Abramov described components he refers to as
\textbf{containers}. A container component is concerned with \textbf{how
things work}.{[}20{]} Containers act as invisible wrappers around
presentational components acting in the sense of UNIX filters. Their job
is to fetch data from child nodes, aggregating events, interacting with
the model and push state back to the presentational components.
Consequently they might contain more JS and less if any HTML markup. We
probably do not need to utilize shadow DOM as no styles are involved.

@@

The \emph{root container}
\texttt{\textless{}shop-checkout\textgreater{}} is basically just an
encapsulation layer in terms of service functionalities. Encapsulation
of CSS will not be an issue at this point as no styling is involved. A
simplified \emph{root container} for the checkout might look like the
following code snippet:

@@

The second layer of the checkout microservice like
\texttt{\textless{}sign-in\textgreater{}} or
\texttt{\textless{}shipping-details\textgreater{}} still contains mostly
logic and little or less styling. The job of a \emph{container
component} is to control its underlying presentational components and to
define and spell a \textbf{set of ACTIONS towards the model}. Every
\emph{container} mounted directly under the service root may have a
dedicated area inside the worker thread reserved for its duties. For
example, a typical \texttt{\textless{}sign-in\textgreater{}} contains
merely two fields, username and password and a submit button. The
\emph{container component} waits for a submit action, aggregating the
credentials, and might add some semantics to them. Field values and the
action message will be dispatched towards the model for further
processing like initiating an authorization process.

@@

\section{5. Outlook}\label{outlook}

In the search for an independent browsernative approach this paper left
out \textbf{third-party libraries} purposely. It has to be pointed out
that the example given in chapter 4 of this paper is far from being
usable in production. At first, from a technical perspective web
components are relatively young and at the time of writing only usable
in \textbf{newer Chrome and Opera browsers}. Most probably it will take
time until native web components can be safely used without additional
legacy support. The troublesome \emph{HTML imports} spec will certainly
need even longer which inevitably opens up the question if web
components are useful from today onwards. And second, do web components
align with the notion of simplicity if too many drawbacks and possible
pitfalls require attention?!

Any profound discussion about web components would be incomplete without
ever touching \textbf{component alternatives} like React or Angular. In
the last couple of years both introduced new concepts to web development
like the idea of state, passing data as properties and declarative event
management just to name a few. Even though they use the notion of
components too, they differentiate much from the philosophy of web
components. Web components are more about the DOM whereas others focus
on JS. Implicitly this approach opens up a case for \textbf{`standards
vs.~opinions'}.

Overall, React's perception about UI development created much of a hype
around the library and technologies like the virtual DOM while web
components have never caught up in momentum at a comparable rate.
Certainly the biggest drawbacks of web components were frequently
changing specs and the uneasiness of using an underperforming polyfilled
version. In the highly competitive field of UI frameworks, \textbf{web
components clearly missed a real selling point} compared to opinionated
frameworks so far. Notable web component projects like Bosonic\footnote{https://github.com/bosonic/bosonic/}
or Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
than alive. For the future it remains unclear if web components
eventually gain wider adoption within the UI developing community or
will be adopted as low-level technology for framework development
emphasized by Sebastian Markbage, one of the React creators.{[}21{]}

Even though the drawbacks seems heavy weight for the moment things will
certainly change for good with flourishing browser adoption. A good
example of recent efforts is the ``reactish'' library Skate\footnote{https://github.com/skatejs/skatejs/}.
Following the functional rendering model from React, Skate combines
native methods with additional functionalities known from React like
declarative event management. Given the focus on native technologies it
only weights around 4 KB (minified and gzipped) which is ten times less
than React. \textbf{Micro UI frameworks} like Skate could certainly be
the near future of web development, offering a real selling point.
Previously mentioned specs could lead to modular, pluggable building
pipelines offering the same features and tooling as React already does.
After all, web components fit better with the microservice
\textbf{decentralization aspect} than `walled garden' frameworks.
